Analysis of $\piup$-$\piup$ lattice QCD scattering data

### 2015-03-17 Paper.tex
Write something about the data generation

% Weighted generation with HMC

% Downloadable ensembles

% Operators on data

% Contractions

>>>>

Monte Carlo methods can be used to simulate quantum systems. Key pieces are
Feynman's path integral formalism and Euclidean spacetime where the metric
tensor $\tens\eta$ has signature $\pm 4$ and the weight factor in the path
integral is $\exp(-S_\text{cl})$ instead of $\exp(\iup S_\text{cl})$ where
$S_\text{cl}$ is the classical action
\parencite[Section~2]{Creutz/Statistical_Approach_QM}. A Monte Carlo algorithm
like the Metropolis algorithm can then be used to generate configurations which
are already distributed by the weight of the classical action. Therefore,
expectational values of observables are simply calculated and averaged over all
available configurations \parencite[(3.7)]{Creutz/Statistical_Approach_QM}.

The group of Andreas Kell, Martin Efferz and Simon Blanke have used this for
the harmonic oscillator in their Computational Physics project this year. I
also did this in my bachelor's thesis: \parencite{Ueding/Bachelorarbeit}.

For quantum chromodynamics (QCD), the theory of the strong force, this
methodology can be used as well. The action of QCD is used with a hybrid Monte
Carlo algorithm to generate configurations of gauge fields on a four
dimensional lattice. This process is done for several ensembles which differ in
parameters like the quark mass. Ensemble generation takes a lot of computing
time, so the results are made available to download.
% TODO Would be nice to have a location where this could be downloaded.

As with the harmonic oscillator on the lattice, expectation values of
observables can be computed for each configuration of the ensembles and then
averaged (without weight) over all ensembles yielding an estimate for the
observable. Since this uses a field theory, particles of interested can be
created at the beginning of time and annihilated at the end of the temporal
lattice extent. This way, specific particle configurations (like
$\piup$–$\piup$ scattering) can be examined without needing to generate new
configurations.

Using the correct operators, it is possible to construct correlation functions
which contain information about the energies/masses of the involved particles,
analogously to \parencite[(4.14)]{Creutz/Statistical_Approach_QM} for the
harmonic oscillator. I was given those correlation functions to analyze.
